---
title: "Course Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible and Research Course

This is first project of reproducible and research course. The project is based on response questions about the dataset "Activity monitoring data". This dataset contains personal activity monitoring device. This device collects data at 5 minute intervals through out the day.


##Loading and preprocessing data
Show any code that is needed to


1. Load the data (i.e. read.csv())
```{r Load data}
library(plyr)
library(reshape2)
filezip <- "MotoringData.zip"

if(!file.exists(filezip))
	{
	fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
      download.file(fileUrl, destfile="MotoringData.zip", method ="auto")
	}

## Unzip the dataset
if (!file.exists("activity.csv")) 
	{ 
  	unzip("MotoringData.zip") 
	}

## read the table

Activity <- read.csv("activity.csv")

```

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r Process}
#Convert to date format
Activity$date <- as.Date(as.character(Activity$date), "%Y-%m-%d")

#Omit NA
Activity <- na.omit(Activity)
```

##What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r NumberStepsDay }
Number_Steps_PerDay <- aggregate(steps ~ date, data = Activity, FUN= sum)
Number_Steps_PerDay
```
```
         date steps
1  2012-10-02   126
2  2012-10-03 11352
3  2012-10-04 12116
4  2012-10-05 13294
5  2012-10-06 15420
6  2012-10-07 11015
7  2012-10-09 12811
8  2012-10-10  9900
9  2012-10-11 10304
10 2012-10-12 17382
11 2012-10-13 12426
12 2012-10-14 15098
13 2012-10-15 10139
14 2012-10-16 15084
15 2012-10-17 13452
16 2012-10-18 10056
17 2012-10-19 11829
18 2012-10-20 10395
19 2012-10-21  8821
20 2012-10-22 13460
21 2012-10-23  8918
22 2012-10-24  8355
23 2012-10-25  2492
24 2012-10-26  6778
25 2012-10-27 10119
26 2012-10-28 11458
27 2012-10-29  5018
28 2012-10-30  9819
29 2012-10-31 15414
30 2012-11-02 10600
31 2012-11-03 10571
32 2012-11-05 10439
33 2012-11-06  8334
34 2012-11-07 12883
35 2012-11-08  3219
36 2012-11-11 12608
37 2012-11-12 10765
38 2012-11-13  7336
39 2012-11-15    41
40 2012-11-16  5441
41 2012-11-17 14339
42 2012-11-18 15110
43 2012-11-19  8841
44 2012-11-20  4472
45 2012-11-21 12787
46 2012-11-22 20427
47 2012-11-23 21194
48 2012-11-24 14478
49 2012-11-25 11834
50 2012-11-26 11162
51 2012-11-27 13646
52 2012-11-28 10183
53 2012-11-29  7047
```


2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
```{r histogram}
#Plot the histogram
hist(Number_Steps_PerDay$steps, col= "gold", main="Total of Steps Per Day", xlab ="Date")


```

3. Calculate and report the mean and median of the total number of steps taken per day
``` {r mediaAndMean}
#Media
Media_Per_Day <- mean(Number_Steps_PerDay$steps)
Media_Per_Day
```
```
[1] 10766.19
```
```{r medianPe}
#Median

Median_Per_Day <- median(Number_Steps_PerDay$steps)
Median_Per_Day
```
```
[1] 10765
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r TimeSeries}

#Average of steps per day

Number_Steps_PerDay_Mean <- aggregate(steps ~ interval , data = Activity, FUN= mean)

#Plot type l
plot(Number_Steps_PerDay_Mean$interval, Number_Steps_PerDay_Mean$steps, type="l", main="The Average of Steps Per Day", ylab ="Number of  Steps", xlab ="Dates")

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

``` {r MaximunSteps}
#Maximun of step per day

positi <- which(Number_Steps_PerDay_Mean$steps== max(Number_Steps_PerDay_Mean$steps))

#Date of maximun

Number_Steps_PerDay_Mean[positi,1]

``` 
``` 
[1] 835

``` 

##Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r nrows}
#Number Total NA

Activity <- read.csv("activity.csv")
NumberNA <- sum(is.na(Activity))
NumberNA

```
```
[1] 2304
```
2. Devise a strategy for filling in all of the missing values in the dataset-

I replaced the NA's with 0.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r replace}
#Replace the NA with 0

Activity <- read.csv("activity.csv")
Data_Dat_withOut_NA <- replace(Activity, is.na(Activity), 0)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r medianReplace}
#Histogram with replace the NA with 0.
Number_Steps_PerDay_OutNA <- aggregate(steps ~ date, data = Data_Dat_withOut_NA, FUN= sum)
hist(Number_Steps_PerDay_OutNA$steps, col= "gold", main="Total of Steps Per Day without NA", xlab ="Date")

#Mean and Median without NA

Number_Steps_PerDay_Mean <- mean(Number_Steps_PerDay_OutNA$steps)
Number_Steps_PerDay_Mean
```
```
[1] 9354.23
```
``` {r MedianOut}
Number_Steps_PerDay_Median <- median(Number_Steps_PerDay_OutNA$steps)
Number_Steps_PerDay_Median
```
```
[1] 10395
```
Clearly, the histograms with NA and with NA equal to 0 are different. In the same way, the media and median. The replace NA with other values affect the final result. 


##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r weekday}

#Merge a column a weekday or weekend
#Convert to date format

Data_Dat_withOut_NA$date <- as.Date(as.character(Data_Dat_withOut_NA$date), "%Y-%m-%d")
weekday <- weekdays(Data_Dat_withOut_NA$date)
Data_Dat_withOut_NA_days <- cbind(Data_Dat_withOut_NA,weekday) 

Data_Dat_withOut_NA_days$weekday <- gsub("lunes|martes|miércoles|jueves|viernes", "weekday", Data_Dat_withOut_NA_days$weekday)
Data_Dat_withOut_NA_days$weekday <- gsub("sábado|domingo", "weekend", Data_Dat_withOut_NA_days$weekday)
#Convert to factor
Data_Dat_withOut_NA_days$weekday <- factor(Data_Dat_withOut_NA_days$weekday)

```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r figures }
library(lattice)
Data_Total <- aggregate (steps ~ interval +weekday,data=Data_Dat_withOut_NA_days, mean)
xyplot(steps ~ interval| weekday, Data_Total,type="l", layout=c(1,2))
```


